Automatic conversion of colloquial Finnish to standard Finnish

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
+1
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58613 72bbbca6-d526-0410-a7d9-f06f51895060

% supervised systems trained on the normalised corpus outperform those trained on the standardised corpus (WHY?)
% could be because of better word alignment

@@

The systems for normalisation are able to improve out-of-vocabulary rates in 
many cases, most likely as the online statistical system that we used is trained
on more formal texts. Frequent contractions such as \emph{onks?} `is there?',
\emph{oo} `not' and \emph{vaa} `only' are found untranslated in the output.

%! ooh mitä onks siellä lunta
%ref: ooh what, is there snow?
%colloq: ooh what onks there snow
%book: what! Is there snow?
%pbmt: ooh what if there's snow
%rules: ooh what if there's snow

%! mä vaa protestoin et ei siinä mitää tavurajaa oo i:den välissä
%ref: I was just protesting that there's no syllable break between the two i's
%colloq: I VAA you do not protest it makes no tavurajaa oo i: in between
%book: I just protest, that's not anything tavurajaa no i in between
%pbmt: I only protest that there is no tavurajaa in between the i-den
%rules: I VAA you do not protest it makes no tavurajaa no i in between

%! (kyse oli siis vaan jostain 3 rivistä loopin sisällä)
% ref: (it was only like 3 lines inside a loop)
% colloq: (it was, therefore, but for some three lines within the loop)
% book: (it was, therefore, only three to a row within a loop)
% pbmt: (it was, therefore, only some three-line within the loop)
% rules: (it was Siisi only some 3 lines of the loop inside)

>>>>

It is interesting to note that supervised systems trained on the normalised section of the corpus outperform
those trained on the standardised corpus. One explanation for this could be that the corpus size is small, so 
that the word alignments are not as reliable on the standardised corpus which is by nature not a word-for-word 
conversion.

@@

The systems for normalisation are able to improve out-of-vocabulary rates in 
many cases, most likely as the online statistical system that we used is trained
on more formal texts. Frequent contractions such as \emph{onks?} `is there?',
\emph{oo} `not' and \emph{vaa} `only' are found untranslated in the output, but 
are easily converted by the systems.

%! ooh mitä onks siellä lunta
%! mitä! Onko siellä lunta?
%ref: ooh what, is there snow?
%colloq: ooh what onks there snow
%book: what! Is there snow?
%pbmt: ooh what if there's snow
%rules: ooh what if there's snow

%! mä vaa protestoin et ei siinä mitää tavurajaa oo i:den välissä
%! minä vain protestoin, että ei siinä mitään tavurajaa ole i:den välissä
%ref: I was just protesting that there's no syllable break between the two i's
%colloq: I VAA you do not protest it makes no tavurajaa oo i: in between
%book: I just protest, that's not anything tavurajaa no i in between
%pbmt: I only protest that there is no tavurajaa in between the i-den
%rules: I VAA you do not protest it makes no tavurajaa no i in between

%! (kyse oli siis vaan jostain 3 rivistä loopin sisällä)
%! (kyse oli siis vain 3:sta rivistä loopin sisällä)
% ref: (it was only like 3 lines inside a loop)
% colloq: (it was, therefore, but for some three lines within the loop)
% book: (it was, therefore, only three to a row within a loop)
% pbmt: (it was, therefore, only some three-line within the loop)
% rules: (it was Siisi only some 3 lines of the loop inside)
